---
title: "ASE startup"
author: Anna L Tyler
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    collapsed: no
    toc: yes
    toc_float: yes
---


## Introduction
The purpose of this workflow is to start looking
at some CC-RIX mice to assess ASE in adults in 
a similar way to pmid:28238550, in which they 
used correlation of allele-specific expression
across F1 mice to identify genes with biased
allele expression in bulk-RNASeq from adult 
tissues.


```{r param}
rm(list = ls())
library(here)
```

```{r load_code}
all.fun <- list.files(here("Code"), full.names = TRUE)
for(i in 1:length(all.fun)){source(all.fun[i])}
```

```{r libraries, message = FALSE, warning = FALSE, error = FALSE}
needed.libraries <- c("pheatmap", "DT")
load_libraries(needed.libraries)
```

I am using the GBRS output that Matt Vincent just made
for an old CGD data set looking at heart and kidneys
in CC-RIX mice. They will not be heterozygous everywhere,
and in fact, some of these mice seem to be F1s of the
same strain, but do have some heterozygosity left. Not
sure what that's about. Maybe these are still early 
CC-RIX?

If GBRS is supposed to genotype from RNASeq, how do
they deal with potential mono-allelic expression?
I guess, for the most part, this will be extremely 
rare, and not be consistent across cells, so there
should always be some of both alleles expressed.

I will start with the kidney data because there
are more samples.

```{r load_data}
sample.dir <- "CGD_CCRIX_Kidney_RNASeq"
file.pattern <- "diploid.genes.expected_read_counts" #calls are identical across animals
#file.pattern <- "diploid.genes.tpm"


sample.sheet <- read.csv(here("Data", sample.dir, "CGD_CCRIX_Kidney_RNASeq.SampleSheet.csv"))
split.strains <- strsplit(gsub(".fastq.gz", "", basename(sample.sheet[,"fastq_1"])), "_")
strains <- sapply(split.strains, function(x) x[1])
u_strain <- unique(strains)

data.files <- sapply(sample.sheet[,1], 
  function(x) list.files(here("Data", sample.dir, "gbrs_analysis", "output", x, "gbrs"), 
  pattern = file.pattern, full.names = TRUE))

ase <- lapply(data.files, function(x) as.matrix(read.delim(x)))
```

There are `r length(u_strain)` unique crosses in this data set:

```{r crosses}
print(u_strain)
```

We will look at each one separately. 
First, we looked at 


we identify all transcripts with heterozygous 
alleles. Then we 

```{r find_hets}


number_calls <- function(call.list){
  geno.cols <- sapply(call.list, function(x) x[,"notes"])
  #flag any positions that have different calls across animals
  u_calls <- apply(geno.cols, 1, unique)
  num.calls <- sapply(u_calls, length)
  prop.table <- (table(num.calls)/nrow(geno.cols))*100

  unique.hap <- apply(geno.cols, 1, function(x) unique(unlist(strsplit(x, ""))))
  num.hap <- sapply(unique.hap, length)

  #par(mfrow = c(1,2))
  barplot_with_num(signif(prop.table, 2), ylab = "Percent",  
    xlab = "Number of different genotypes")

  hist(num.hap, xlab = "Haplotypes", main = "")
}

find_hets <- function(call.list, min.count = 10){
  

  geno.cols <- sapply(call.list, function(x) x[,"notes"])
  
  #filter to sites for which all animals have the same call
  #we'll call these high-confidence sites, although, we should
  #investigate why some sites have different calls.
  same.call <- apply(geno.cols, 1, function(x) all(x == x[1]))
  all.same <- length(which(same.call))/nrow(geno.cols)

  sub.geno <- geno.cols[which(same.call),]
  sub.counts <- lapply(call.list, function(x) apply(x[which(same.call), LETTERS[1:8]], 2, as.numeric))
  #name rows with gene names
  for(i in 1:length(sub.counts)){
    rownames(sub.counts[[i]]) <- call.list[[i]][which(same.call),1]
  }

  #filter out genes for which there was no call for any 
  has.call <- which(sub.geno[,1] != "None")
  filtered.geno <- sub.geno[has.call,]
  filtered.counts <- lapply(sub.counts, function(x) x[has.call,])

  #find heterozygous sites
  unique.hap <- apply(filtered.geno, 1, function(x) unique(unlist(strsplit(x, ""))))
  num.hap <- sapply(unique.hap, length)
  #hist(num.hap)
  het.idx <- which(num.hap == 2)
  #length(het.idx)

  het.counts <- lapply(filtered.counts, function(x) x[het.idx,])

  #convert to a list of matrices with counts for all alleles
  allele.counts <- lapply(1:nrow(het.counts[[1]]), function(x) sapply(het.counts, function(y) t(y[x,])))
  names(allele.counts) <- rownames(het.counts[[1]])
  #rownames disappeared, add them back
  for(i in 1:length(allele.counts)){
    rownames(allele.counts[[i]]) <- LETTERS[1:8]
  }

  #remove genes with counts that are too los
  count.sum <- sapply(allele.counts, function(x) log(sum(x)+1))
  #hist(count.sum)
  keep.genes <- which(count.sum > min.count)

  has.counts <- allele.counts[keep.genes]

  final.counts <- lapply(has.counts, function(x) x[which(rowSums(x) > 0),])
  #WTF, these are all identical. There's no variation at all! I don't think
  #expected counts is what I want...
  return(final.counts)  
}
```

I first just looked at how many different genotype calls
there were across supposedly isogenic animals. There are
an alarmingly high number of different calls. For three
of the crosses there are more sites with two different 
calls than with only one. Why is this?

```{r count_hets, fig.height = 4, fig.width = 8}
for(i in 1:length(u_strain)){
  par(mfrow = c(1,2))
  cross.idx <- which(strains == u_strain[i])
  number_calls(call.list = ase[cross.idx])
  mtext(u_strain[i], side = 3, line = -1.5, outer = TRUE)
}
```


For now, I guess I'll restrict my analysis to 
only sites for which all animals have the same 
call.


```{r get_hets}
cross_hets <- vector(mode = "list", length = length(u_strain))
names(cross_hets) <- u_strain
for(i in 1:length(u_strain)){
  cross.idx <- which(strains == u_strain[i])
  cross_hets[[i]] <- find_hets(call.list = ase[cross.idx])
}
```

What the heck! These numbers look so strange. I don't understand
what all of this stuff is, but it is not what I need. 