---
title: "GTEx startup"
author: Anna L Tyler
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    collapsed: no
    toc: yes
    toc_float: yes
---


I'm going to try looking at the GTEx data here.
Just more exploration. 



```{r param}
rm(list = ls())
library(here)
```

```{r load_code}
all.fun <- list.files(here("Code"), full.names = TRUE)
for(i in 1:length(all.fun)){source(all.fun[i])}
```

```{r libraries, message = FALSE, warning = FALSE, error = FALSE}
needed.libraries <- c("pheatmap", "DT", "gprofiler2", "vioplot", 
  "VGAM", "parallel", "doParallel", "igraph")
load_libraries(needed.libraries)
```

There is just one big expression matrix. Here we parse 
out which individuals and tissues each sample belongs 
to. I've also downloaded the meta data for GTEx V8.

We filtered out genes with low read counts (<10) and 
expressed too few tissues (<3).


```{r load_data}
#This chunk takes a little while the first time through

sample.att <- as.matrix(read.delim(here("Data", "GTEx", 
  "GTEx_Analysis_v8_Annotations_SampleAttributesDS.txt")))
ind.att <- as.matrix(read.delim(here("Data", "GTEx", 
  "GTEx_Analysis_v8_Annotations_SubjectPhenotypesDS.txt")))

age.group <- sapply(strsplit(ind.att[,"AGE"], "-"), function(x) round(mean(as.numeric(x))))
ind.att <- cbind(ind.att, age.group)
ind.sex <- ind.att[,"SEX"]
ind.sex[which(ind.att[,"SEX"] == 1)] <- "Male"
ind.sex[which(ind.att[,"SEX"] == 2)] <- "Female"
ind.att <- cbind(ind.att, "sex" = ind.sex)

expr.file <- here("Data", "GTEx", "expression_individual.RDS")
gene.info.file <- here("Data", "GTEx", "info_gene.RDS")
sample.info.file <- here("Data", "GTEx", "info_sample.RDS")
if(!file.exists(expr.file)){
  #reading in this file takes several minutes
  expr <- read.delim(here("Data", "GTEx", "phASER_WASP_GTEx_v8_matrix.gw_phased.txt"))
  expr.info <- expr[,1:4]
  saveRDS(expr.info, gene.info.file)
  
  hap.levels <- expr[,5:ncol(expr)]

  sample.id <- gsub(".", "-", colnames(hap.levels), fixed = TRUE)
  ind.id <- sapply(strsplit(sample.id, "-"), function(x) paste(x[1], x[2], sep = "-"))
  samp.idx <- sapply(sample.id, function(x) which(sample.att[,1] == x))
  matched.sample.info <- sample.att[samp.idx,c("SAMPID", "SMTSD")]
  samp.info <- cbind(ind.id, matched.sample.info)
  saveRDS(samp.info, sample.info.file)

  #group expression by individual
  u_ind <- unique(ind.id)
  #length(u_ind)
  ind_expr <- lapply(u_ind, function(x) hap.levels[,which(samp.info[,1] == x),drop=FALSE])
  names(ind_expr) <- u_ind

  #split each individual into two matrices for the two haplotypes
  #and filter down to genes that are expressed in at least three
  #tissues with at least 10 reads in each tissue.
  #this takes about half an hour on a laptop. Go make some tea.
  hap.expr <- vector(mode = "list", length = length(ind_expr))
  names(hap.expr) <- u_ind
  for(i in 1:length(ind_expr)){
    report.progress(i, length(ind_expr))
    hap1 <- apply(ind_expr[[i]], 2, 
      function(x) as.numeric(sapply(strsplit(x, "|", fixed = TRUE), function(y) y[1])))
    hap2 <- apply(ind_expr[[i]], 2, 
      function(x) as.numeric(sapply(strsplit(x, "|", fixed = TRUE), function(y) y[2])))
    rownames(hap1) <- rownames(hap2) <- sapply(strsplit(expr.info[,"name"], ".", fixed = TRUE), function(x) x[1])

    #replace sample IDs with tissue names
    samp.id <- colnames(hap1) 
    tissue.id <- samp.info[match(gsub(".", "-", samp.id, fixed = TRUE), samp.info[,"SAMPID"]), "SMTSD"]
    colnames(hap1) <- colnames(hap2) <- tissue.id

    enough.tissues1 <- which(apply(hap1, 1, function(x) length(which(x > 0))) >= 3)
    enough.tissues2 <- which(apply(hap2, 1, function(x) length(which(x > 0))) >= 3)
    enough.tx.idx <- union(enough.tissues1, enough.tissues2)

    if(length(enough.tx.idx) == 0){
      next()
    }

    sub.hap1 <- hap1[enough.tx.idx,]
    sub.hap2 <- hap2[enough.tx.idx,]

    enough.counts1 <- which(rowSums(sub.hap1) >= 10)
    enough.counts2 <- which(rowSums(sub.hap2) >= 10)
    enough.counts.idx <- union(enough.counts1, enough.counts2)

    if(length(enough.counts.idx) == 0){
      next()
    }

    filtered.hap1 <- sub.hap1[enough.counts.idx,]
    filtered.hap2 <- sub.hap2[enough.counts.idx,]

    #head(filtered.hap1)
    #head(filtered.hap2)
    #idx = 2;cbind(filtered.hap1[idx,], filtered.hap2[idx,])
    #comp_binom(cbind(filtered.hap1[idx,], filtered.hap2[idx,]))

    hap.list <- list("haplotype1" = filtered.hap1, "haplotype2" = filtered.hap2)
    hap.expr[[i]] <- hap.list
  }
  #some individuals have fewer than 3 tissues sampled
  #make sure we only include with enough tissues
  has.vals <- which(sapply(hap.expr, length) > 0)
  saveRDS(hap.expr[has.vals], expr.file)
}else{
  hap.expr <- readRDS(expr.file)
  gene.loc.info <- readRDS(gene.info.file)
}

gene.name.file <- here("Data", "GTEx", "gene_info.RDS")
if(!file.exists(gene.name.file)){
  all_gene <- unique(unlist(lapply(hap.expr, function(x) rownames(x[[1]]))))
  split.gene <- sapply(strsplit(all_gene, ".", fixed = TRUE), function(x) x[[1]])
  #gene.info <- gconvert(split.gene, organism = "hsapiens")
  gene.info <- get_gene_info(geneID = split.gene, IDtype = "ensembl_gene_id",
    organism = "human")
  saveRDS(gene.info, gene.name.file)
}else{
  gene.info <- readRDS(gene.name.file)
}
```




```{r plot_fun}
gene_name <- function(gene.id = NULL, gene.name = NULL){
  if(is.null(gene.id)){
    gene.id <- gene.info[which(gene.info[,"external_gene_name"] == gene.name)[1],"ensembl_gene_id"]
    if(length(gene.id) == 0){
      stop(paste("I can't find the gene", gene.name))
    }
  }
  if(is.null(gene.name)){
    gene.name <- gene.info[which(gene.info[,"ensembl_gene_id"] == gene.id)[1],"external_gene_name"]
    if(length(gene.name) == 0){
      stop(paste("I can't find the gene", gene.id))
    }
  }
  result <- c("name" = as.character(gene.name), "id" = as.character(gene.id))
} 

comp_binom <- function(haplotype.table){

  #remove rows with all zeros or nas
  not.na <- which(!is.na(rowSums(haplotype.table)))
  sub.table <- haplotype.table[not.na,]
  #high.table <- sub.table[which(rowSums(sub.table) >= 10),]
  #plot(high.table)
  #ratios <- high.table[,1]/rowSums(high.table)
  #hist(ratios)
  #total.count <- rowSums(high.table)
  #plot(total.count+1, ratios, log = "xy")
  
  fit.bb <- try(vglm(sub.table ~ 1, family=betabinomial), silent = TRUE)
  if(class(fit.bb) == "try-error"){
    return(NA)
  }
  fit.b <- vglm(sub.table~1, binomialff)
  loglik.bb <- logLik(fit.bb)
  loglik.b <- logLik(fit.b)

  lambda_lr <- (loglik.b - loglik.bb)*-2
  p <- 1-pchisq(lambda_lr, df = 1)
  return(p)
}

test_gene_fit <- function(gene.name = NULL, gene.id = NULL, tissue.name = NULL){

  gene.info <- gene_name(gene.id, gene.name)
  gene.name <- gene.info["name"]
  gene.id <- gene.info["id"]
  
  #get the expression across tissues for each individual
  hap1.expr <- lapply(hap.expr, function(x) x[[1]][which(rownames(x[[1]]) == gene.id),])
  hap2.expr <- lapply(hap.expr, function(x) x[[2]][which(rownames(x[[2]]) == gene.id),])

  if(!is.null(tissue.name)){
    hap1.tissue <- lapply(hap1.expr, function(x) x[grep(tissue.name, names(x), ignore.case = TRUE),drop=FALSE])
    hap2.tissue <- lapply(hap2.expr, function(x) x[grep(tissue.name, names(x), ignore.case = TRUE),drop=FALSE])
    plot.label <- paste(gene.name, tissue.name)
  }else{
    hap1.tissue <- hap1.expr
    hap2.tissue <- hap2.expr
    plot.label <- paste(gene.name, "all tissues")
  }

  has.vals <- which(sapply(hap1.tissue, length) > 0)
  paired.haps <- lapply(has.vals, function(x) cbind(hap1.tissue[[x]], hap2.tissue[[x]]))
  all.hap1 <- Reduce("c", lapply(paired.haps, function(x) x[,1]))
  all.hap2 <- Reduce("c", lapply(paired.haps, function(x) x[,2]))

  fit.p <- sapply(paired.haps, comp_binom)
  return(fit.p)
}

#tests all genes for a given individual
test_ind_genes <- function(ind_expr, tissue = NULL, verbose = FALSE){
  hap1 <- ind_expr[[1]]
  hap2 <- ind_expr[[2]]
  if(!is.null(tissue)){
    tissue.idx <- grep(tissue, colnames(hap1), ignore.case = TRUE)
    if(length(tissue.idx) <= 5){
      return(NA)
    }
  }
  if(verbose){
    gene.p <- unlist(lapply_pb(1:nrow(hap1), function(x) comp_binom(cbind(hap1[x,], hap2[x,]))))
  }else{
    gene.p <- sapply(1:nrow(hap1), function(x) comp_binom(cbind(hap1[x,], hap2[x,])))
  }
  names(gene.p) <- rownames(hap1)
  gene.p <- gene.p[which(!is.na(gene.p))]
  return(gene.p)
}


plot_gene_cor <- function(gene.name = NULL, gene.id = NULL, 
  tissue.name = NULL, plot.results = TRUE){

  gene.info <- gene_name(gene.id, gene.name)
  gene.name <- gene.info["name"]
  gene.id <- gene.info["id"]

  #get the expression across tissues for each individual
  hap1.expr <- lapply(hap.expr, function(x) x[[1]][which(rownames(x[[1]]) == gene.id),])
  hap2.expr <- lapply(hap.expr, function(x) x[[2]][which(rownames(x[[2]]) == gene.id),])

  if(!is.null(tissue.name)){
    hap1.tissue <- lapply(hap1.expr, function(x) x[grep(tissue.name, names(x), ignore.case = TRUE),drop=FALSE])
    hap2.tissue <- lapply(hap2.expr, function(x) x[grep(tissue.name, names(x), ignore.case = TRUE),drop=FALSE])
    plot.label <- paste(gene.name, tissue.name)
  }else{
    hap1.tissue <- hap1.expr
    hap2.tissue <- hap2.expr
    plot.label <- paste(gene.name, "all tissues")
  }

  has.vals <- which(sapply(hap1.tissue, length) > 0)
  paired.haps <- lapply(has.vals, function(x) cbind(hap1.tissue[[x]], hap2.tissue[[x]]))
  all.hap1 <- Reduce("c", lapply(paired.haps, function(x) x[,1]))
  all.hap2 <- Reduce("c", lapply(paired.haps, function(x) x[,2]))

  #test.check <- sapply(paired.haps, comp_binom)
  #qqunif.plot(test.check)
  
  hap.cor <- sapply(paired.haps, function(x) cor(x[,1], x[,2], method = "spearman"))
  hap.ratios <- lapply(paired.haps, function(x) x[,1]/rowSums(x))
  
  if(plot.results){
    par(mfrow = c(1,3))
    plot.with.model(all.hap1, all.hap2, xlab = "Haplotype 1", ylab = "Haplotype 2", 
      pch = 16, main = gene.name, report = "cor.test")
    hist(hap.cor, xlab = "Spearman correlation", breaks = 50, 
      main = "Haplotype Correlation", xlim = c(-1, 1))
    abline(v = 0, col = "red")
    mtext(plot.label, side = 3, outer = TRUE, line = -1.5, cex = 1.2, font = 2)
    hist(unlist(hap.ratios), main = "Allele Ratios", xlab = "ratio", xlim = c(0, 1))
  }

  result <- list("correlations" = hap.cor, "haplotype_counts" = paired.haps,
    "ratios" = hap.ratios)
  invisible(result)
}

get_ind_expr <- function(gene.name = NULL, gene.id = NULL, 
  ind.name, plot.results = FALSE){

  gene.info <- gene_name(gene.id, gene.name)
  gene.name <- gene.info["name"]
  gene.id <- gene.info["id"]
  
  ind.idx <- which(names(hap.expr)== ind.name)
  if(length(ind.idx) == 0){
    stop(paste("I can't find the individual", ind.name))
  }

  ind.expr <- hap.expr[[ind.idx]]
  gene.idx <- which(rownames(ind.expr[[1]]) == gene.id)
  
  #get the expression across tissues for each individual
  hap1.expr <- ind.expr[[1]][which(rownames(ind.expr[[1]]) == gene.id),]
  hap2.expr <- ind.expr[[2]][which(rownames(ind.expr[[2]]) == gene.id),]

  paired.haps <- cbind(hap1.expr, hap2.expr)

  if(plot.results){
    xlim <- c(min(hap1.expr, na.rm = TRUE), (max(hap1.expr, na.rm = TRUE)*1.1))
    plot(hap1.expr, hap2.expr, xlab = "Haplotype 1", ylab = "Haplotype 2",
      pch = 16, main = gene.name, xlim = xlim)    
    text(hap1.expr, hap2.expr, labels = names(hap1.expr), cex = 0.6, 
      pos = 4)
    abline(0,1)
  }

  return(paired.haps)
}
```


```{r ind_gene_test, eval = FALSE}


imprinted.genes <- c("FAM50B", "PXDC1", "GRB10", "PLAGL1", "HYMAI")
gene.cor <- lapply(imprinted.genes, plot_gene_cor)
names(gene.cor) <- imprinted.genes

par(mfrow = c(1,2))
boxplot(lapply(gene.cor, function(x) x[[1]]), main = "Correlations")
abline(h = c(0, 0.5))
boxplot(lapply(gene.cor, function(x) x[[2]]), main = "Ratios")
abline(h = 1)

imprinted <- plot_gene_cor(gene.name = "HYMAI")
imprinted.p <- test_gene_fit("HYMAI")
imprinted.fdr <- p.adjust(imprinted.p, method = "BH")
qqunif.plot(imprinted.fdr)
which(imprinted.fdr <= 0.1)

plot(-log10(imprinted.fdr), -log10(biallele.fdr))


biallele <- plot_gene_cor("GAB1")
biallele.p <- test_gene_fit("GAB1")

biallele.fdr <- p.adjust(biallele.p, method = "BH")
qqunif.plot(biallele.fdr)
which(biallele.fdr <= 0.1)


imprinted$haplotype_counts[which(imprinted.fdr <= 0.1)]
biallele$haplotype_counts[which(biallele.fdr <= 0.1)]

test.mat <- biallele$haplotype_counts[which(biallele.fdr <= 0.1)][[1]]
pheatmap(test.mat)
comp_binom(test.mat)

par(mfrow = c(1,2))
qqunif.plot(biallele.p, plot.label = "biallelic")
qqunif.plot(imprinted.p, plot.label = "imprinted")

test <- plot_gene_cor("KCNK9")
plot_gene_cor("KCNK9", tissue.name = "brain")

plot_gene_cor("KL")
plot_gene_cor("KL", tissue.name = "brain")
```


```{r plot_haps}

plot_expr_as_mat <- function(cor_result, plot.label = ""){
  hap.result <- cor_result$haplotype_counts
  ind.idx <- match(names(hap.result), ind.att[,"SUBJID"])
  
  ind.sex <- as.factor(ind.att[ind.idx,"sex"])
  ind.age.group <- as.numeric(ind.att[ind.idx,"age.group"])
  annot.df <- data.frame("age.group" = rep(ind.age.group, each = 2), 
    "sex" = rep(ind.sex, each = 2))
  rownames(annot.df) <- paste(rep(names(hap.result), each = 2), c(1,2), sep = "_")
  
  all.tx <- unique(unlist(lapply(hap.result, rownames)))

  all.expr <- vector(mode = "list", length = length(hap.result))
  names(all.expr) <- names(hap.result)

  for(i in 1:length(hap.result)){
    ind.mat <- matrix(NA, nrow = length(all.tx), ncol = 2)
    rownames(ind.mat) <- all.tx
    tx <- rownames(hap.result[[i]])
    ind.mat[tx,] <- hap.result[[i]]
    all.expr[[i]] <- ind.mat
  }

  ind.order <- order(sapply(all.expr, function(x) mean(x, na.rm = TRUE)))
  expr.mat <- Reduce("cbind", all.expr[ind.order])
  colnames(expr.mat) <- rownames(annot.df)
  row.order <- order(rowMeans(expr.mat, na.rm = TRUE))  

  pheatmap(log10(expr.mat[row.order,]+1), annotation_col = annot.df,
    cluster_rows = FALSE, cluster_cols = FALSE, main = plot.label, show_colnames = FALSE)
}


plot_deviations <- function(cor_result, main = "", log.transform = TRUE, 
  plot.results = TRUE, ratio.thresh = 0.9){
  #collect all haplotypes

  all.haps <- cbind(unlist(lapply(cor_result$haplotype_counts, function(x) x[,1])),
    unlist(lapply(cor_result$haplotype_counts, function(x) x[,2])))
  if(log.transform){
    all.haps <- log10(all.haps+1)
  }
  #plot(log10(all.haps+1))
  ratios <- all.haps[,1]/rowSums(all.haps)
  above.thresh <- which(ratios >= ratio.thresh)
  below.thresh <- which(ratios <= 1-ratio.thresh)

  dev.haps <- all.haps[c(names(above.thresh), names(below.thresh)),]

  above.ind <- Reduce("rbind", strsplit(names(above.thresh), ".", fixed = TRUE))
  below.ind <- Reduce("rbind", strsplit(names(below.thresh), ".", fixed = TRUE)) 
  skewed.ratios <- rbind(above.ind, below.ind)
  ordered.table <- skewed.ratios[order(skewed.ratios[,1]),]
  rownames(ordered.table) <- NULL

  if(plot.results){
    layout.mat <- matrix(c(1,2,3,4), nrow = 2, byrow = FALSE)
    layout(layout.mat, widths = c(0.8, 1))

    pt.col <- rep("black", nrow(all.haps))
    pt.col[above.thresh] <- "red"
    pt.col[below.thresh] <- "blue"
    plot(all.haps[,1], all.haps[,2], xlab = "Haplotype 1",
      ylab = "Haplotype 2", col = pt.col, main = main, pch = 16)
    abline(0,1)
    
    hist(ratios, main = "ratios", xlab = "ratios")
    abline(v = c(ratio.thresh, 1-ratio.thresh))

    #pheatmap(log10(dev.haps+1))
    par(mar = c(4, 16, 0, 2))
    imageWithText(log10(dev.haps+1), show.text = FALSE, 
      use.pheatmap.colors = TRUE, row.text.shift = 0.5,
      row.text.cex = 0.7)

    par(mar = c(4, 16, 0, 2))
    barplot(sort(ratios[c(above.thresh, below.thresh)]), 
      las = 2, horiz = TRUE, cex.names = 0.7,
      xlab = "Allele Ratio")
    plot.dim <- par("usr")
    segments(x0 = 0.5, y0 = -1, y1 = plot.dim[4]-1)

  }
  invisible(ordered.table)
}
```

```{r ind_tests, eval = FALSE}
gene.name = "KL"
test <- plot_gene_cor(gene.name)
#quartz(width = 11, height = 6)
deviants <- plot_deviations(test, main = gene.name, ratio.thresh = 0.9, 
  log.transform = TRUE)


#quartz(width = 11, height = 6)
show.max <- min(20, nrow(deviants))
par(mfrow = c(1,2), mar = c(4,10,4,2))
barplot(rev(sort(table(deviants[,1]), decreasing = TRUE)[1:show.max]), 
  las = 2, horiz = TRUE, main = "Individuals with skewed ratios", 
  xlab = "Tissues", cex.names = 0.8)
par(mar = c(4,10,4,2))
barplot(rev(sort(table(deviants[,2]), decreasing = TRUE)[1:show.max]), 
  las = 2, horiz = TRUE, cex.names = 0.8,
  main = "Tissues with skewed ratios", xlab = "Individuals")
mtext(gene.name, outer = TRUE, side = 3, line = -2.5, font = 2, cex = 1.2)

plot_expr_as_mat(test, gene.name)

high.ind <- sort(table(deviants[,1]), decreasing = TRUE)
all.ind.p <- sapply(names(high.ind), function(x) comp_binom(get_ind_expr(gene.name, x)))
sig.ind <- which(all.ind.p < 0.00001)

par(mfrow = c(3,3))
for(i in 1:length(sig.ind)){
  test.ind <- get_ind_expr(gene.name, names(sig.ind)[i])
  plot(log10(test.ind), pch = 16, xlim = c(0, 3))
  text(log10(test.ind[,1]), log10(test.ind[,2]), labels = rownames(test.ind), 
    pos = 4, cex = 0.5)
  abline(0,1)
}


test.ratio <- test.ind[,1]/rowSums(test.ind)
test.ratio <- test.ratio[which(!is.na(test.ratio))]
ratio.order <- order(test.ratio)
barplot(test.ratio[ratio.order], las = 2)
imageWithText(log10(test.ind[ratio.order,]+1), use.pheatmap.colors = TRUE, 
  show.text = FALSE, row.text.shift = 0.5)

test.age <- as.numeric(ind.att[match(names(test$correlations), 
  ind.att[,"SUBJID"]), "age.group"])
mean.expr <- sapply(test$haplotype_counts, function(x) mean(log10(x+1)))

expr.by.age <- lapply(unique(test.age), function(x) mean.expr[which(test.age == x)])
names(expr.by.age) <- unique(test.age)

cor.by.age <- lapply(unique(test.age), function(x) test$correlations[which(test.age == x)])
names(cor.by.age) <- unique(test.age)


par(mfrow = c(1,2))
vioplot(expr.by.age, main = "Expression by age", col = "lightgray")
mtext("age group", side = 1, line = 2.5)
mtext("log expression", side = 2, line = 2.5)
stripchart(expr.by.age, vertical = TRUE, method = "jitter", pch = 16, 
  col = "#dd1c77", add = TRUE, cex = 0.5)

vioplot(cor.by.age, main = "Haplotype correlation by age", col = "lightgray")
mtext("age group", side = 1, line = 2.5)
mtext("Spearman correlation", side = 2, line = 2.5)
stripchart(cor.by.age, vertical = TRUE, method = "jitter", pch = 16, 
  col = "#dd1c77", add = TRUE, cex = 0.5)

mtext(gene.name, side = 3, outer = TRUE, line = -1.5, cex = 1.2, font = 2)
```


```{r a_few_more, eval = FALSE}
gene.name <- c("MAPK1", "ARL17A", "PARD6G")
test.results <- vector(mode = "list", length = length(gene.name))
names(test.results)  <- gene.name

pdf("~/Desktop/check_genes.pdf", width = 8, height = 4)
for(i in 1:length(gene.name)){
  report.progress(i, length(gene.name))
  test.results[[i]] <- plot_gene_cor(gene.name[i])
}
dev.off()

```

the stats tell us if a given gene has biased

allele expression in a single individual (gene x individual)
  how many genes in each individual have biased expression?
  in how many individuals does each gene have biased expression?


genes x tissues x individuals

for each gene
  count how many individuals have skew in each tissue
  

for each individual count how many genes are skewed in each tissue
for each tissue count 

```{r testing, eval = FALSE}
test <- test_ind_genes(hap.expr[[1]])
```

```{r comp_all}

all_p_file <- here("Results", "processed_data", "Individual_Gene_P_values.RDS")

if(!file.exists(all_p_file)){
  # Parallel computation using 4 cores
  #cl <- makeCluster(detectCores()-1)
  #cl <- makeCluster(4)
  cl <- makeCluster(20)
  registerDoParallel(cl)
  clusterExport(cl, c("test_ind_genes", "comp_binom", "vglm", "betabinomial", 
    "binomialff", "logLik", "pchisq", "try"))
  #all_p <- foreach(ind_expr = hap.expr[1:4]) %dopar% test_ind_genes(ind_expr)
  all_p <- foreach(ind_expr = hap.expr) %dopar% test_ind_genes(ind_expr)
  stopCluster(cl)
  saveRDS(all_p, all_p_file)
}else{
  all_p <- readRDS(all_p_file)
}
```

```{r just_brain, eval = FALSE}

brain_p_file <- here("Results", "processed_data", "Individual_Gene_P_values_brain.RDS")

if(!file.exists(brain_p_file)){
  # Parallel computation using 4 cores
  #cl <- makeCluster(detectCores()-1)
  #cl <- makeCluster(4)
  cl <- makeCluster(20)
  registerDoParallel(cl)
  clusterExport(cl, c("test_ind_genes", "comp_binom", "vglm", "betabinomial", 
    "binomialff", "logLik", "pchisq", "try"))
  #all_p <- foreach(ind_expr = hap.expr[1:4]) %dopar% test_ind_genes(ind_expr)
  brain_p <- foreach(ind_expr = hap.expr) %dopar% test_ind_genes(ind_expr)
  stopCluster(cl)
  saveRDS(all_p, all_p_file)
}else{
  all_p <- readRDS(all_p_file)
}
```


```{r alpha_thresh}
alpha.thresh <- 1e-6
```

## Number of biased genes per individual

There were many genes, even at a very conservative,
threshold that had significantly biased gene expression
(adjusted p < `r alpha.thresh`).
The histogram below shows the distribution of biased
genes per individual.

Most people had fewer than 1000 biased genes, but 
some had many more. 

```{r process_p}
adj.p <- lapply(all_p, function(x) p.adjust(x, "BH"))
sig.p <- lapply(adj.p, function(x) which(x <= alpha.thresh))
names(adj.p) <- names(sig.p) <- names(hap.expr)
num.sig <- sapply(sig.p, length)
hist(num.sig, breaks = 50, 
  main = "Number of biased genes per individual", xlab = "Number of genes")
```

## Cause of the inter-individual variation

The main driver of the number of genes identified
as having biased expression, seems to be the number
of tissues that were measured. Having more tissues
measured gives you greater power to detect biased
gene expression. The plot below shows how the number
of biased genes changes as a function of the number
of tissues measured. The relationship is very strong,
although there are also some crazy outliers.

This means that in order to look at inter-individual
variation in the number of biased genes, we'll need
to do some different statistics. We could re-sample
at a lower rate, but this will result in a lot of
false negatives. What about an adaptive p value 
threshold that takes into account the number of 
tissues measured in each person? That seems as if 
it risks a lot of false positives...

Could we do some combination of those things? 

```{r bias_by_ind}
num.tissues <- sapply(hap.expr, function(x) ncol(x[[1]]))
boxplot(num.sig~num.tissues,
  ylab = "Number of Biased Genes", xlab = "Number of Tissues")
model <- lm(num.sig~num.tissues)
abline(model, lwd = 2, col = "#de2d26")


#brain.idx <- sapply(hap.expr, function(x) grep("brain", colnames(x[[1]]), ignore.case = TRUE))
#num.brain <- sapply(brain.idx, length)
#hist(num.brain)
```

## Adaptive p values

Below we look at what kinds of thresholds it would 
take to even out the number of biased genes per individual.

```{r num_detect}
med.num <- round(median(sapply(num.sig, median)))
```

The plot below shows at what p value cutoff you can detect
`r med.num` genes (which is the median number of genes
detected across all patients at p = `r alpha.thresh`) 
as a function of the number of tissues sampled.

The horizontal line shows the cutoff p = 0.01.

```{r adaptive_p}

find_thresh <- function(pV, target.num){
  sorted.p <- sort(pV, decreasing = FALSE)
  #plot(-log10(sorted.p+1e-17))
  #abline(v = target.num)
  target.p <- sorted.p[round(target.num)]
  return(target.p)
}

adaptive.thresh <- sapply(adj.p, function(x) find_thresh(x, med.num))
#hist(-log10(adaptive.thresh+1e-17), breaks = 50)


plot.with.model(num.tissues, -log10(adaptive.thresh+1e-17),
  xlab = "Number of tissues", ylab = "-log10(p)", 
  main = paste("Threshold to detect", med.num, "genes"))
abline(h = -log10(0.01))

```

What if we look at the p value distributions for a 
given gene? 
```{r p_by_gene}
tissue.num <- sapply(hap.expr, function(x) ncol(x[[1]]))

#gene.name = "MAPK1"
#gene.name <- sample(gene.info[,"external_gene_name"], 1)

#get gene ID
gene.id.info <- gene_name(gene.name = gene.name)
#find p value indices for the gene
gene.idx <- sapply(adj.p, function(x) which(names(x) == gene.id.info["id"]))
has.vals <- which(sapply(gene.idx, length) > 0)
gene.p <- unlist(sapply(adj.p, function(x) x[which(names(x) == gene.id.info["id"])]))

#numbers of subjects in which the gene was detected and had a significant p value
detected.in <- length(gene.p)
sig.in <- length(which(gene.p <= 1e-6))

#collect all haplotype counts for this gene
all.gene.haps <- lapply(1:length(gene.idx), 
  function(x) if(length(gene.idx[[x]]) > 0){cbind(hap.expr[[x]][[1]][gene.idx[[x]],], hap.expr[[x]][[2]][gene.idx[[x]],])})

#get the average counts for this gene across all tissues in each subject
avg.count <- sapply(all.gene.haps, function(x) if(length(x) > 0){mean(x)}else{NA})
#hist(avg.count)
#low.count <- which(avg.count < 1)
#all.gene.haps[[low.count[1]]]

#make a table of all haplotype counts together
hap.table <- Reduce("rbind", all.gene.haps)
high.haps <- hap.table[which(rowSums(hap.table) >= 10),]


if(detected.in > 10){
  par(mfrow = c(2,2))

  #qq plot for p values
  qqunif.plot(test);abline(h = 6)

  #plot of p values vs. number of tissues in each subj.
  plot.with.model(tissue.num[has.vals], -log10(gene.p+1e-17), xlab = "number of tissues",
    ylab = "-log10(p)")

  #haplotypes counted against each other
  plot(high.haps, xlab = "Haplotype 1", ylab = "Haplotype 2",
    main = gene.name)
  abline(0,1)

  #see where the significant p values are coming from.
  #tissues with low counts, or high counts
  plot(avg.count[which(!is.na(avg.count))], 
    -log10(test+1e-17), xlab = "Avg read count across tissues",
    ylab = "-log10(p)", pch = 16)

  #plot.text(paste("Found in", detected.in, 
  #  "people\nSignificant in", sig.in,
  #  "people:\n", signif((sig.in/detected.in)*100, 2), "percent"))

    mtext(gene.name, outer = TRUE, side = 3, line = -1.5, font = 2, cex = 1.2)
}
```

Also, each of these people has different tissues measured,
which could affect which genes are detected, but I think 
we probably can't worry about that. 

```{r max_p}
max.p <- 0.01
```

## Implementing adaptive p value

The plot below shows the number of significantly biased
genes, if we use the p value threshold derived from the
number of tissues sampled. We set a maximum adjusted 
p value threshold at `r max.p`.

There is no longer a relationship between the number of 
significantly biased genes and the number of tissues
sampled. I'm not sure if this is valid...

```{r set_thresh}
model <- lm(-log10(adaptive.thresh+1e-17)~num.tissues)
coefs <- coef(model)
thresh.by.tissue <- sapply(1:40, function(x) (coefs[2]*x)+coefs[1])
p.by.tissue <- 10^(thresh.by.tissue*-1)
p.by.tissue[which(p.by.tissue > max.p)] <- max.p
#set a threshold for each individual based on the number
#of tissues they had measured
ind.thresh <- sapply(num.tissues, function(x) p.by.tissue[x])

new.sig <- lapply(1:length(adj.p), function(x) which(adj.p[[x]] <= ind.thresh[x]))
names(new.sig) <- names(sig.p)
num.new.sig <- sapply(new.sig, length)
#hist(num.new.sig, breaks = 50)

plot.with.model(num.tissues, num.new.sig, xlab = "Number of tissues",
ylab = "Number of significantly biased genes")
```

## No relationship with bias and sex or age

The plots below show that there is no relationship
between the number of significantly biased genes and
age or sex. I'm showing the log10 of significantly
biased genes, because some individuals have huge
numbers.

```{r age_bias}
ind.idx <- match(names(new.sig), ind.att[,1])
vioplot(log10(num.new.sig+1)~as.numeric(ind.att[ind.idx,"age.group"]),
  xlab = "age group", ylab = "log10 Number of significantly biased genes",
  main = "No difference in biased gene count by age")

vioplot(log10(num.new.sig+1)~as.factor(ind.att[ind.idx,"sex"]),
  xlab = "sex", ylab = "Number of significantly biased genes",
  main = "No difference in biased gene count by sex")
```

## Number of shared genes

The histogram below shows how many individuals have 
biased expression in each gene. The vast majority
are only biased in a single individual.

```{r common_genes}
u_genes <- unique(unlist(lapply(new.sig, names)))
gene.chr <- gene.info[match(u_genes, gene.info[,"ensembl_gene_id"]),"chromosome_name"]

inc.mat <- matrix(0, nrow = length(new.sig), ncol = length(u_genes))
rownames(inc.mat) <- names(new.sig)
colnames(inc.mat) <- u_genes
for(i in 1:length(new.sig)){
  inc.mat[i,names(new.sig[[i]])] <- 1
}
genes.shared <- colSums(inc.mat)
hist(genes.shared, breaks = 50, xlab = "Number of individuals",
  main = 'Number of individuals sharing n biased genes')
```

Not surprisingly, this distribution is close to a power
law, but not quite.

```{r power}
num.shared <- 1:max(genes.shared)
shared.by <- sapply(num.shared, function(x) length(which(genes.shared == x)))
#plot(num.shared, shared.by, log = "xy")
plot(log(num.shared+1), log(shared.by+1), 
  xlab = "log number of individuals sharing the gene",
  ylab = "log number of genes")

test <- lm(log(shared.by+1)~log(num.shared+1))
#summary(test)
abline(test)
```

The number of genes shared by at least n people is
shown below. The horizontal line shows y = 100.

```{r shared_genes}
#collect all genes shared by each number of individuals
shared.genes <- lapply(1:max(num.shared), function(x) colnames(inc.mat)[which(genes.shared == x)])
names(shared.genes) <- 1:max(num.shared)


#make lists of genes with at least x number of individuals sharing biased expression
count.shared <- 10:500
at.least.some <- lapply(count.shared, function(x) unlist(shared.genes[x:length(shared.genes)]))
at.least.num <- sapply(at.least.some, length)
names(at.least.num) <- count.shared
plot(count.shared, at.least.num, xlab = "Number of individuals",
  ylab = "Number of genes")
abline(h = 100)
```

## Enrichments for shared genes {.tabset .tabset-fade .tabset-pills}

The enrichment plots below show functional enrichments
for biased genes shared by at least n people. There
are a lot of extracellular enrichments in the less
shared genes, and a lot of immunological enrichments
in the more shared genes.

```{r enrich_num, fig.width = 7, fig.height = 7, results = "asis"}
for(i in seq(200, 500, 100)){
  cat("###", i, "genes\n")
  enrich <- gost(at.least.some[[which(count.shared == i)]], 
    organism = "hsapiens", sources = c("GO", "KEGG", "HPA"))
  plot.enrichment(enrich, num.terms = 30, 
    plot.label = paste("Biased genes shared by at least", i, "people"))
  cat("\n\n")
}
```


```{r nets, eval = FALSE}
genes.which <- at.least.some[[which(count.shared == 200)]]

#pheatmap(inc.mat[,genes.which], show_rownames = FALSE)
plot.decomp(inc.mat[,genes.which], 
  color.by = ind.att[match(rownames(inc.mat), ind.att[,1]),"sex",drop=FALSE]) #dots are individuals
  
plot.decomp(t(inc.mat[,genes.which])) #dots are genes

inc.net <- graph_from_incidence_matrix(inc.mat[,genes.which])
inc.proj <- bipartite_projection(inc.net)

ind.proj <- inc.proj[[1]]
gene.proj <- inc.proj[[2]]

degree(ind.proj)
degree(gene.proj)

```

```{r enrich}
#too many! Even 10 takes a long time
#ind.enrich <- lapply(lapply(new.sig, names)[1:10], function(x) gost(x, organism = "hsapiens"))
#plot.enrichment.group(ind.enrich)

```
