---
title: "GTEx startup"
author: Anna L Tyler
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    collapsed: no
    toc: yes
    toc_float: yes
---


I'm going to try looking at the GTEx data here.
Just more exploration. 



```{r param}
rm(list = ls())
library(here)
```

```{r load_code}
all.fun <- list.files(here("Code"), full.names = TRUE)
for(i in 1:length(all.fun)){source(all.fun[i])}
```

```{r libraries, message = FALSE, warning = FALSE, error = FALSE}
needed.libraries <- c("pheatmap", "DT", "gprofiler2", "vioplot", 
  "VGAM", "parallel", "doParallel")
load_libraries(needed.libraries)
```

There is just one big expression matrix, and I'll need 
to parse out which individuals and tissues each sample
belongs to. I've also downloaded the meta data for GTEx
V8

```{r load_data}
#This chunk takes a little while the first time through

sample.att <- as.matrix(read.delim(here("Data", "GTEx", 
  "GTEx_Analysis_v8_Annotations_SampleAttributesDS.txt")))
ind.att <- as.matrix(read.delim(here("Data", "GTEx", 
  "GTEx_Analysis_v8_Annotations_SubjectPhenotypesDS.txt")))

age.group <- sapply(strsplit(ind.att[,"AGE"], "-"), function(x) round(mean(as.numeric(x))))
ind.att <- cbind(ind.att, age.group)
ind.sex <- ind.att[,"SEX"]
ind.sex[which(ind.att[,"SEX"] == 1)] <- "Male"
ind.sex[which(ind.att[,"SEX"] == 2)] <- "Female"
ind.att <- cbind(ind.att, "sex" = ind.sex)

expr.file <- here("Data", "GTEx", "expression_individual.RDS")
gene.info.file <- here("Data", "GTEx", "info_gene.RDS")
sample.info.file <- here("Data", "GTEx", "info_sample.RDS")
if(!file.exists(expr.file)){
  #reading in this file takes several minutes
  expr <- read.delim(here("Data", "GTEx", "phASER_WASP_GTEx_v8_matrix.gw_phased.txt"))
  expr.info <- expr[,1:4]
  saveRDS(expr.info, gene.info.file)
  
  hap.levels <- expr[,5:ncol(expr)]

  sample.id <- gsub(".", "-", colnames(hap.levels), fixed = TRUE)
  ind.id <- sapply(strsplit(sample.id, "-"), function(x) paste(x[1], x[2], sep = "-"))
  samp.idx <- sapply(sample.id, function(x) which(sample.att[,1] == x))
  matched.sample.info <- sample.att[samp.idx,c("SAMPID", "SMTSD")]
  samp.info <- cbind(ind.id, matched.sample.info)
  saveRDS(samp.info, sample.info.file)

  #group expression by individual
  u_ind <- unique(ind.id)
  #length(u_ind)
  ind_expr <- lapply(u_ind, function(x) hap.levels[,which(samp.info[,1] == x),drop=FALSE])
  names(ind_expr) <- u_ind

  #split each individual into two matrices for the two haplotypes
  #and filter down to genes that are expressed.
  #this takes about half an hour on a laptop. Go make some tea.
  hap.expr <- vector(mode = "list", length = length(ind_expr))
  names(hap.expr) <- u_ind
  for(i in 1:length(ind_expr)){
    report.progress(i, length(ind_expr))
    hap1 <- apply(ind_expr[[i]], 2, 
      function(x) as.numeric(sapply(strsplit(x, "|", fixed = TRUE), function(y) y[1])))
    hap2 <- apply(ind_expr[[i]], 2, 
      function(x) as.numeric(sapply(strsplit(x, "|", fixed = TRUE), function(y) y[2])))
    rownames(hap1) <- rownames(hap2) <- sapply(strsplit(expr.info[,"name"], ".", fixed = TRUE), function(x) x[1])

    #replace sample IDs with tissue names
    samp.id <- colnames(hap1) 
    tissue.id <- samp.info[match(gsub(".", "-", samp.id, fixed = TRUE), samp.info[,"SAMPID"]), "SMTSD"]
    colnames(hap1) <- colnames(hap2) <- tissue.id

    has.vals1 <- which(rowSums(hap1) > 0)
    has.vals2 <- which(rowSums(hap2) > 0)
    has.vals <- union(has.vals1, has.vals2)
    sub.hap1 <- hap1[has.vals,,drop=FALSE]
    sub.hap2 <- hap2[has.vals,,drop=FALSE]
    hap.list <- list("haplotype1" = sub.hap1, "haplotype2" = sub.hap2)
    hap.expr[[i]] <- hap.list
  }
  saveRDS(hap.expr, expr.file)
}else{
  hap.expr <- readRDS(expr.file)
}



gene.name.file <- here("Data", "GTEx", "gene_name_info.RDS")
if(!file.exists(gene.name.file)){
  all_gene <- unique(unlist(lapply(hap.expr, function(x) rownames(x[[1]]))))
  split.gene <- sapply(strsplit(all_gene, ".", fixed = TRUE), function(x) x[[1]])
  gene.name.info <- gconvert(split.gene, organism = "hsapiens")
  saveRDS(gene.name.info, gene.name.file)
}else{
  gene.name.info <- readRDS(gene.name.file)
}
```

```{r plot_fun}
gene_name <- function(gene.id = NULL, gene.name = NULL){
  if(is.null(gene.id)){
    gene.id <- gene.name.info[which(gene.name.info[,"name"] == gene.name),"target"]
    if(length(gene.id) == 0){
      stop(paste("I can't find the gene", gene.name))
    }
  }
  if(is.null(gene.name)){
    gene.name <- gene.name.info[which(gene.name.info[,"target"] == gene.id),"name"]
    if(length(gene.name) == 0){
      stop(paste("I can't find the gene", gene.id))
    }
  }
  result <- c("name" = as.character(gene.name), "id" = as.character(gene.id))
} 

comp_binom <- function(haplotype.table){

  #remove rows with all zeros or nas
  not.na <- which(!is.na(rowSums(haplotype.table)))
  sub.table <- haplotype.table[not.na,]
  #plot(sub.table)
  
  fit.bb <- try(vglm(sub.table ~ 1, family=betabinomial), silent = TRUE)
  if(class(fit.bb) == "try-error"){
    return(NA)
  }
  fit.b <- vglm(sub.table~1, binomialff)
  loglik.bb <- logLik(fit.bb)
  loglik.b <- logLik(fit.b)

  lambda_lr <- (loglik.b - loglik.bb)*-2
  p <- 1-pchisq(lambda_lr, df = 1)
  return(p)
}

test_gene_fit <- function(gene.name = NULL, gene.id = NULL, tissue.name = NULL){

  gene.info <- gene_name(gene.id, gene.name)
  gene.name <- gene.info["name"]
  gene.id <- gene.info["id"]
  
  #get the expression across tissues for each individual
  hap1.expr <- lapply(hap.expr, function(x) x[[1]][which(rownames(x[[1]]) == gene.id),])
  hap2.expr <- lapply(hap.expr, function(x) x[[2]][which(rownames(x[[2]]) == gene.id),])

  if(!is.null(tissue.name)){
    hap1.tissue <- lapply(hap1.expr, function(x) x[grep(tissue.name, names(x), ignore.case = TRUE),drop=FALSE])
    hap2.tissue <- lapply(hap2.expr, function(x) x[grep(tissue.name, names(x), ignore.case = TRUE),drop=FALSE])
    plot.label <- paste(gene.name, tissue.name)
  }else{
    hap1.tissue <- hap1.expr
    hap2.tissue <- hap2.expr
    plot.label <- paste(gene.name, "all tissues")
  }

  has.vals <- which(sapply(hap1.tissue, length) > 0)
  paired.haps <- lapply(has.vals, function(x) cbind(hap1.tissue[[x]], hap2.tissue[[x]]))
  all.hap1 <- Reduce("c", lapply(paired.haps, function(x) x[,1]))
  all.hap2 <- Reduce("c", lapply(paired.haps, function(x) x[,2]))

  fit.p <- sapply(paired.haps, comp_binom)
  return(fit.p)
}

#tests all genes for a given individual
test_ind_genes <- function(ind_expr, verbose = FALSE){
  hap1 <- ind_expr[[1]]
  hap2 <- ind_expr[[2]]
  if(verbose){
    gene.p <- unlist(lapply_pb(1:nrow(hap1), function(x) comp_binom(cbind(hap1[x,], hap2[x,]))))
  }else{
    gene.p <- sapply(1:nrow(hap1), function(x) comp_binom(cbind(hap1[x,], hap2[x,])))
  }
  names(gene.p) <- rownames(hap1)
  gene.p <- gene.p[which(!is.na(gene.p))]
  return(gene.p)
}


plot_gene_cor <- function(gene.name = NULL, gene.id = NULL, 
  tissue.name = NULL, plot.results = TRUE){

  gene.info <- gene_name(gene.id, gene.name)
  gene.name <- gene.info["name"]
  gene.id <- gene.info["id"]

  #get the expression across tissues for each individual
  hap1.expr <- lapply(hap.expr, function(x) x[[1]][which(rownames(x[[1]]) == gene.id),])
  hap2.expr <- lapply(hap.expr, function(x) x[[2]][which(rownames(x[[2]]) == gene.id),])

  if(!is.null(tissue.name)){
    hap1.tissue <- lapply(hap1.expr, function(x) x[grep(tissue.name, names(x), ignore.case = TRUE),drop=FALSE])
    hap2.tissue <- lapply(hap2.expr, function(x) x[grep(tissue.name, names(x), ignore.case = TRUE),drop=FALSE])
    plot.label <- paste(gene.name, tissue.name)
  }else{
    hap1.tissue <- hap1.expr
    hap2.tissue <- hap2.expr
    plot.label <- paste(gene.name, "all tissues")
  }

  has.vals <- which(sapply(hap1.tissue, length) > 0)
  paired.haps <- lapply(has.vals, function(x) cbind(hap1.tissue[[x]], hap2.tissue[[x]]))
  all.hap1 <- Reduce("c", lapply(paired.haps, function(x) x[,1]))
  all.hap2 <- Reduce("c", lapply(paired.haps, function(x) x[,2]))

  #test.check <- sapply(paired.haps, comp_binom)
  #qqunif.plot(test.check)
  
  hap.cor <- sapply(paired.haps, function(x) cor(x[,1], x[,2], method = "spearman"))
  hap.ratios <- lapply(paired.haps, function(x) x[,1]/rowSums(x))
  
  if(plot.results){
    par(mfrow = c(1,3))
    plot.with.model(all.hap1, all.hap2, xlab = "Haplotype 1", ylab = "Haplotype 2", 
      pch = 16, main = gene.name, report = "cor.test")
    hist(hap.cor, xlab = "Spearman correlation", breaks = 50, 
      main = "Haplotype Correlation", xlim = c(-1, 1))
    abline(v = 0, col = "red")
    mtext(plot.label, side = 3, outer = TRUE, line = -1.5, cex = 1.2, font = 2)
    hist(unlist(hap.ratios), main = "Allele Ratios", xlab = "ratio", xlim = c(0, 1))
  }

  result <- list("correlations" = hap.cor, "haplotype_counts" = paired.haps,
    "ratios" = hap.ratios)
  invisible(result)
}

get_ind_expr <- function(gene.name = NULL, gene.id = NULL, 
  ind.name, plot.results = FALSE){

  gene.info <- gene_name(gene.id, gene.name)
  gene.name <- gene.info["name"]
  gene.id <- gene.info["id"]
  
  ind.idx <- which(names(hap.expr)== ind.name)
  if(length(ind.idx) == 0){
    stop(paste("I can't find the individual", ind.name))
  }

  ind.expr <- hap.expr[[ind.idx]]
  gene.idx <- which(rownames(ind.expr[[1]]) == gene.id)
  
  #get the expression across tissues for each individual
  hap1.expr <- ind.expr[[1]][which(rownames(ind.expr[[1]]) == gene.id),]
  hap2.expr <- ind.expr[[2]][which(rownames(ind.expr[[2]]) == gene.id),]

  paired.haps <- cbind(hap1.expr, hap2.expr)

  if(plot.results){
    xlim <- c(min(hap1.expr, na.rm = TRUE), (max(hap1.expr, na.rm = TRUE)*1.1))
    plot(hap1.expr, hap2.expr, xlab = "Haplotype 1", ylab = "Haplotype 2",
      pch = 16, main = gene.name, xlim = xlim)    
    text(hap1.expr, hap2.expr, labels = names(hap1.expr), cex = 0.6, 
      pos = 4)
    abline(0,1)
  }

  return(paired.haps)
}
```


```{r ind_gene_test, eval = FALSE}


imprinted.genes <- c("FAM50B", "PXDC1", "GRB10", "PLAGL1", "HYMAI")
gene.cor <- lapply(imprinted.genes, plot_gene_cor)
names(gene.cor) <- imprinted.genes

par(mfrow = c(1,2))
boxplot(lapply(gene.cor, function(x) x[[1]]), main = "Correlations")
abline(h = c(0, 0.5))
boxplot(lapply(gene.cor, function(x) x[[2]]), main = "Ratios")
abline(h = 1)

imprinted <- plot_gene_cor(gene.name = "HYMAI")
imprinted.p <- test_gene_fit("HYMAI")
imprinted.fdr <- p.adjust(imprinted.p, method = "BH")
qqunif.plot(imprinted.fdr)
which(imprinted.fdr <= 0.1)

plot(-log10(imprinted.fdr), -log10(biallele.fdr))


biallele <- plot_gene_cor("GAB1")
biallele.p <- test_gene_fit("GAB1")

biallele.fdr <- p.adjust(biallele.p, method = "BH")
qqunif.plot(biallele.fdr)
which(biallele.fdr <= 0.1)


imprinted$haplotype_counts[which(imprinted.fdr <= 0.1)]
biallele$haplotype_counts[which(biallele.fdr <= 0.1)]

test.mat <- biallele$haplotype_counts[which(biallele.fdr <= 0.1)][[1]]
pheatmap(test.mat)
comp_binom(test.mat)

par(mfrow = c(1,2))
qqunif.plot(biallele.p, plot.label = "biallelic")
qqunif.plot(imprinted.p, plot.label = "imprinted")

test <- plot_gene_cor("KCNK9")
plot_gene_cor("KCNK9", tissue.name = "brain")

plot_gene_cor("KL")
plot_gene_cor("KL", tissue.name = "brain")
```


```{r plot_haps}

plot_expr_as_mat <- function(cor_result, plot.label = ""){
  hap.result <- cor_result$haplotype_counts
  ind.idx <- match(names(hap.result), ind.att[,"SUBJID"])
  
  ind.sex <- as.factor(ind.att[ind.idx,"sex"])
  ind.age.group <- as.numeric(ind.att[ind.idx,"age.group"])
  annot.df <- data.frame("age.group" = rep(ind.age.group, each = 2), 
    "sex" = rep(ind.sex, each = 2))
  rownames(annot.df) <- paste(rep(names(hap.result), each = 2), c(1,2), sep = "_")
  
  all.tx <- unique(unlist(lapply(hap.result, rownames)))

  all.expr <- vector(mode = "list", length = length(hap.result))
  names(all.expr) <- names(hap.result)

  for(i in 1:length(hap.result)){
    ind.mat <- matrix(NA, nrow = length(all.tx), ncol = 2)
    rownames(ind.mat) <- all.tx
    tx <- rownames(hap.result[[i]])
    ind.mat[tx,] <- hap.result[[i]]
    all.expr[[i]] <- ind.mat
  }

  ind.order <- order(sapply(all.expr, function(x) mean(x, na.rm = TRUE)))
  expr.mat <- Reduce("cbind", all.expr[ind.order])
  colnames(expr.mat) <- rownames(annot.df)
  row.order <- order(rowMeans(expr.mat, na.rm = TRUE))  

  pheatmap(log10(expr.mat[row.order,]+1), annotation_col = annot.df,
    cluster_rows = FALSE, cluster_cols = FALSE, main = plot.label, show_colnames = FALSE)
}


plot_deviations <- function(cor_result, main = "", log.transform = TRUE, 
  plot.results = TRUE, ratio.thresh = 0.9){
  #collect all haplotypes

  all.haps <- cbind(unlist(lapply(cor_result$haplotype_counts, function(x) x[,1])),
    unlist(lapply(cor_result$haplotype_counts, function(x) x[,2])))
  if(log.transform){
    all.haps <- log10(all.haps+1)
  }
  #plot(log10(all.haps+1))
  ratios <- all.haps[,1]/rowSums(all.haps)
  above.thresh <- which(ratios >= ratio.thresh)
  below.thresh <- which(ratios <= 1-ratio.thresh)

  dev.haps <- all.haps[c(names(above.thresh), names(below.thresh)),]

  above.ind <- Reduce("rbind", strsplit(names(above.thresh), ".", fixed = TRUE))
  below.ind <- Reduce("rbind", strsplit(names(below.thresh), ".", fixed = TRUE)) 
  skewed.ratios <- rbind(above.ind, below.ind)
  ordered.table <- skewed.ratios[order(skewed.ratios[,1]),]
  rownames(ordered.table) <- NULL

  if(plot.results){
    layout.mat <- matrix(c(1,2,3,4), nrow = 2, byrow = FALSE)
    layout(layout.mat, widths = c(0.8, 1))

    pt.col <- rep("black", nrow(all.haps))
    pt.col[above.thresh] <- "red"
    pt.col[below.thresh] <- "blue"
    plot(all.haps[,1], all.haps[,2], xlab = "Haplotype 1",
      ylab = "Haplotype 2", col = pt.col, main = main, pch = 16)
    abline(0,1)
    
    hist(ratios, main = "ratios", xlab = "ratios")
    abline(v = c(ratio.thresh, 1-ratio.thresh))

    #pheatmap(log10(dev.haps+1))
    par(mar = c(4, 16, 0, 2))
    imageWithText(log10(dev.haps+1), show.text = FALSE, 
      use.pheatmap.colors = TRUE, row.text.shift = 0.5,
      row.text.cex = 0.7)

    par(mar = c(4, 16, 0, 2))
    barplot(sort(ratios[c(above.thresh, below.thresh)]), 
      las = 2, horiz = TRUE, cex.names = 0.7,
      xlab = "Allele Ratio")
    plot.dim <- par("usr")
    segments(x0 = 0.5, y0 = -1, y1 = plot.dim[4]-1)

  }
  invisible(ordered.table)
}
```

```{r ind_tests, eval = FALSE}
gene.name = "KL"
#gene.name <- sample(gene.name.info[which(gene.name.info[,"name"] != "None"),"name"], 1)
#test <- plot_gene_cor(gene.name, "brain")
#quartz(width = 11, height = 4)
test <- plot_gene_cor(gene.name)
#quartz(width = 11, height = 6)
deviants <- plot_deviations(test, main = gene.name, ratio.thresh = 0.9, 
  log.transform = TRUE)


#quartz(width = 11, height = 6)
show.max <- min(20, nrow(deviants))
par(mfrow = c(1,2), mar = c(4,10,4,2))
barplot(rev(sort(table(deviants[,1]), decreasing = TRUE)[1:show.max]), 
  las = 2, horiz = TRUE, main = "Individuals with skewed ratios", 
  xlab = "Tissues", cex.names = 0.8)
par(mar = c(4,10,4,2))
barplot(rev(sort(table(deviants[,2]), decreasing = TRUE)[1:show.max]), 
  las = 2, horiz = TRUE, cex.names = 0.8,
  main = "Tissues with skewed ratios", xlab = "Individuals")
mtext(gene.name, outer = TRUE, side = 3, line = -2.5, font = 2, cex = 1.2)

plot_expr_as_mat(test, gene.name)

high.ind <- sort(table(deviants[,1]), decreasing = TRUE)
all.ind.p <- sapply(names(high.ind), function(x) comp_binom(get_ind_expr(gene.name, x)))
sig.ind <- which(all.ind.p < 0.00001)

par(mfrow = c(3,3))
for(i in 1:length(sig.ind)){
  test.ind <- get_ind_expr(gene.name, names(sig.ind)[i])
  plot(log10(test.ind), pch = 16, xlim = c(0, 3))
  text(log10(test.ind[,1]), log10(test.ind[,2]), labels = rownames(test.ind), 
    pos = 4, cex = 0.5)
  abline(0,1)
}


test.ratio <- test.ind[,1]/rowSums(test.ind)
test.ratio <- test.ratio[which(!is.na(test.ratio))]
ratio.order <- order(test.ratio)
barplot(test.ratio[ratio.order], las = 2)
imageWithText(log10(test.ind[ratio.order,]+1), use.pheatmap.colors = TRUE, 
  show.text = FALSE, row.text.shift = 0.5)

test.age <- as.numeric(ind.att[match(names(test$correlations), 
  ind.att[,"SUBJID"]), "age.group"])
mean.expr <- sapply(test$haplotype_counts, function(x) mean(log10(x+1)))

expr.by.age <- lapply(unique(test.age), function(x) mean.expr[which(test.age == x)])
names(expr.by.age) <- unique(test.age)

cor.by.age <- lapply(unique(test.age), function(x) test$correlations[which(test.age == x)])
names(cor.by.age) <- unique(test.age)


par(mfrow = c(1,2))
vioplot(expr.by.age, main = "Expression by age", col = "lightgray")
mtext("age group", side = 1, line = 2.5)
mtext("log expression", side = 2, line = 2.5)
stripchart(expr.by.age, vertical = TRUE, method = "jitter", pch = 16, 
  col = "#dd1c77", add = TRUE, cex = 0.5)

vioplot(cor.by.age, main = "Haplotype correlation by age", col = "lightgray")
mtext("age group", side = 1, line = 2.5)
mtext("Spearman correlation", side = 2, line = 2.5)
stripchart(cor.by.age, vertical = TRUE, method = "jitter", pch = 16, 
  col = "#dd1c77", add = TRUE, cex = 0.5)

mtext(gene.name, side = 3, outer = TRUE, line = -1.5, cex = 1.2, font = 2)
```


```{r a_few_more, eval = FALSE}
gene.name <- c("MAPK1", "ARL17A", "PARD6G")
#gene.name <- sample(gene.name.info[which(gene.name.info[,"name"] != "None"),"name"], 100)
test.results <- vector(mode = "list", length = length(gene.name))
names(test.results)  <- gene.name

pdf("~/Desktop/check_genes.pdf", width = 8, height = 4)
for(i in 1:length(gene.name)){
  report.progress(i, length(gene.name))
  test.results[[i]] <- plot_gene_cor(gene.name[i])
}
dev.off()

```

the stats tell us if a given gene has biased

allele expression in a single individual (gene x individual)
  how many genes in each individual have biased expression?
  in how many individuals does each gene have biased expression?


genes x tissues x individuals

for each gene
  count how many individuals have skew in each tissue
  

for each individual count how many genes are skewed in each tissue
for each tissue count 


```{r comp_all}

all_p_file <- here("Results", "processed_data", "Individual_Gene_P_values.RDS")

if(!file.exists(all_p_file)){
  # Parallel computation using 4 cores
  #cl <- makeCluster(detectCores()-1)
  cl <- makeCluster(4)
  #cl <- makeCluster(20)
  registerDoParallel(cl)
  clusterExport(cl, c("test_ind_genes", "comp_binom", "vglm", "betabinomial", 
    "binomialff", "logLik", "pchisq", "try"))
  all_p <- foreach(ind_expr = hap.expr[1:4]) %dopar% test_ind_genes(ind_expr)
  #all_p <- foreach(ind_expr = hap.expr) %dopar% test_ind_genes(ind_expr)
  stopCluster(cl)
  saveRDS(all_p, all_p_file)
}else{
  all_p <- readRDS(all_p_file)
}
```


```{r process_p, eval = FALSE}
adj.p <- lapply(all_p, function(x) p.adjust(x, "BH"))
sig.p <- lapply(adj.p, function(x) which(x <= 1e-6))
#names(sig.p) <- names(hap.expr)[1:4]
#sapply(sig.p, length)
plotVenn(sig.p)
sim.mat <- jaccard.matrix(sig.p)
diag(sim.mat) <- NA
pheatmap(sim.mat)

ind.idx = 1; gene.idx <- 5
gene.info <- gene_name(gene.name = NULL, gene.id = names(sig.p[[ind.idx]])[gene.idx])
par(mfrow = c(1,1))
ind.haps <- get_ind_expr(gene.id = gene.info["id"], 
  ind.name = names(sig.p)[ind.idx], plot.results = TRUE)
cor.result <- plot_gene_cor(gene.name = gene.info["name"])
plot_expr_as_mat(cor.result)
plot_deviations(cor.result, main = gene.info["name"])

u_genes <- unique(unlist(lapply(sig.p, names)))
sig.count <- colSums(sapply(u_genes, function(x) sapply(sig.p, function(y) length(which(names(y) == x)))))
barplot(sig.count)

high.sig <- which(sig.count > 3)
gene.test <- test_gene_fit(gene.id = names(high.sig)[1], tissue.name = NULL)
qqunif.plot(gene.test)
length(which(gene.test <= 1e-6))
length(which(is.na(gene.test)))
```

recalculate p values for just brain 

Questions for when p values are done.
For each gene, how many people have sig p values (taking into account
  which individuals actually had that gene measured)

  
